Containers and R: the Rockerverse and beyond

### 2019-07-30 manuscript.Rmd
distribute content from "further images" section, add @karthik with holepunch
also add some more project links

---
title: "Containers and R: the Rockerverse and beyond"
author:
  - name: Daniel Nüst
    affiliation: Institute for Geoinformatics, University of Münster 
    address:
    - Heisenbergstr. 2, 48149 Münster, Germany
    email: daniel.nuest@uni-muenster.de
  - name: Carl Boettiger
    #affiliation: -
    #email: -
  - name: Robrecht Cannoodt
    #affiliatin: -
    #email: -
  - name: Dirk Eddelbuettel
    #affiliation: -
    #email: -
  - name: Mark Edmondson
    #affiliation: -
    #email: -
  - name: Colin Fay
    #affiliation: -
    #email: -
  - name: Ben Marwick
    #affiliation: -
    #email: -
  - name: Noam Ross
    #affiliation: -
    #email: -
  - name: Nan Xiao
    #affiliation: -
    #email: - 
abstract: >
  The Rocker project provides widely-used Docker images for R across different application scenarios.
  This articles surveys describe downstream projects building upon Rocker.
  We also look beyond Rocker to other projects connecting containerisation with R.
  These use cases and the diversity of applications demonstrate the power of Rocker and containerisation for collaboration, effectivity, scalability, and transparency.
preamble: >
   \usepackage{longtable}
output: rticles::rjournal_article
---

@@

- [MRO](https://github.com/nuest/mro-docker)
- [Renjin](https://github.com/nuest/renjin-docker)
- [pqR](https://github.com/nuest/pqr-docker)
- [FastR](https://github.com/nuest/fastr-docker)

@@

## Further Use-case Docker images [NN]

- part of Rocker (if not mentioned in \citet{RJ-2017-065}) ?
  - Binder
  - GPU [@noamross, @cboettig]
- [Bioconductor](http://bioconductor.org/help/docker/), https://github.com/Bioconductor/bioc_docker
- Gigantum stack
- ...

@@

- `containerit`
- `dockertest`
- `liftr` for OS-level reproducibility [@nanx]
- [`rize` for Shiny](https://github.com/cole-brokamp/rize)
- [`dockerfiler`](https://github.com/ColinFay/dockerfiler)

@@

- R-Hub
- DevOps
  - https://www.opencpu.org/posts/opencpu-with-docker/
- `r-ci`: https://github.com/ColinFay/r-ci
- [dynwrap](https://github.com/dynverse/dynwrap_containers/blob/master/.travis.yml) [@rcannood]
  - For this project, we use travis-ci to build rocker-derived containers, test them, and only push them to docker hub (from travis-ci.org) if the integration tests succeed.
- ...

## Common work environments [@noamross]

- ...
- in education
  - ready to use learning environments
  - [`r-db`](https://github.com/ColinFay/r-db) [@ColinFay]
- [RCloud](https://github.com/att/rcloud/tree/master/docker) ?

## Deployment, processing, cloud

@@

## R versions [@ColinFay]

- run code with Versions of R: https://srv.colinfay.me/r-online
- old R versions: https://github.com/rocker-org/rocker-versioned/issues/138
- [semantic version tags for Rocker](https://github.com/rocker-org/rocker-versioned/#version-tags)
- debugging memory problems across multiple versions: https://github.com/wch/r-debug

### Cross-R testing

Show how you can run the same script in multiple R variants in containers (new contribution).
Ideally the images should be created using `containerit` and controlled with `harbor`/`docker`, i.e. via an R script.

>>>>

---
title: "Containers and R: the Rockerverse and beyond"
author:
  - name: Daniel Nüst
    affiliation: Institute for Geoinformatics, University of Münster 
    address:
    - Heisenbergstr. 2, 48149 Münster, Germany
    email: daniel.nuest@uni-muenster.de
  - name: Carl Boettiger
    #affiliation: -
    #email: -
  - name: Robrecht Cannoodt
    #affiliatin: -
    #email: -
  - name: Dirk Eddelbuettel
    #affiliation: -
    #email: -
  - name: Mark Edmondson
    #affiliation: -
    #email: -
  - name: Colin Fay
    #affiliation: -
    #email: -
  - name: Ben Marwick
    #affiliation: -
    #email: -
  - name: Karthik Ram
    #affiliation: -
    #email: -
  - name: Noam Ross
    #affiliation: -
    #email: -
  - name: Nan Xiao
    #affiliation: -
    #email: - 
abstract: >
  The Rocker project provides widely-used Docker images for R across different application scenarios.
  This articles surveys describe downstream projects building upon Rocker.
  We also look beyond Rocker to other projects connecting containerisation with R.
  These use cases and the diversity of applications demonstrate the power of Rocker and containerisation for collaboration, effectivity, scalability, and transparency.
preamble: >
   \usepackage{longtable}
output: rticles::rjournal_article
---

@@

- [MRO](https://github.com/nuest/mro-docker)
- [Renjin](https://github.com/nuest/renjin-docker)
- [pqR](https://github.com/nuest/pqr-docker)
- [FastR](https://github.com/nuest/fastr-docker)
- [Bioconductor](http://bioconductor.org/help/docker/), https://github.com/Bioconductor/bioc_docker

@@

@@

- [`containerit`](https://github.com/o2r-project/containerit/)
- [`dockertest`](https://github.com/traitecoevo/dockertest/)
- [`liftr`](https://liftr.me/) for OS-level reproducibility [@nanx]
- [`rize` for Shiny](https://github.com/cole-brokamp/rize)
- [`dockerfiler`](https://github.com/ColinFay/dockerfiler)

@@

- R-Hub
- DevOps
  - https://www.opencpu.org/posts/opencpu-with-docker/
- `r-ci`: https://github.com/ColinFay/r-ci
- [dynwrap](https://github.com/dynverse/dynwrap_containers/blob/master/.travis.yml) [@rcannood]
  - For this project, we use travis-ci to build rocker-derived containers, test them, and only push them to docker hub (from travis-ci.org) if the integration tests succeed.

## Common or public work environments [div]

- Binder
  - [`holepunch`](https://github.com/karthik/holepunch) streamlines making an R research compendium Binder-ready [@karthik]
  - [`repo2docker`](https://repo2docker.readthedocs.io/en/latest/config_files.html) (`install.R`, `DESCRPTION`, `runtime.txt`) based on the Rocker images [@nuest]
- GPU [@noamross, @cboettig]
- Gigantum stack
- in education (ready to use teaching/learning environments)
  - [`r-db`](https://github.com/ColinFay/r-db) [@ColinFay]
- [RCloud](https://github.com/att/rcloud/tree/master/docker) ?

## Deployment, processing, cloud [div]

@@

## R versions [@ColinFay, @nuest]

- run code with Versions of R: https://srv.colinfay.me/r-online
- old R versions: https://github.com/rocker-org/rocker-versioned/issues/138
- [semantic version tags for Rocker](https://github.com/rocker-org/rocker-versioned/#version-tags)
- debugging memory problems across multiple versions: https://github.com/wch/r-debug
- cross-R testing: Show how you can run the same script in multiple R variants in containers (new contribution). Ideally the images should be created using `containerit` and controlled with `stevedore`/`docker`, i.e. via an R script.
